MLH1 Results Outline

### 2020-04-22 doc/Results_v2.rmd
transfering code to setup file

@@

```{r female.dif, echo=FALSE}

#how much greater are the high strains?
G_female <-  female_mouse.avs_for.analysis %>% filter(strain == "G")
mean.G.female <- mean(G_female$mean_co, na.rm=TRUE)

LEW_female <-  female_mouse.avs_for.analysis %>% filter(strain == "LEW")
mean.LEW.female <- mean(LEW_female$mean_co, na.rm=TRUE)

#how much greater is MSM female
MSM_female <-  female_mouse.avs_for.analysis %>% filter(strain == "MSM")
mean.MSM.female <- mean(MSM_female$mean_co, na.rm=TRUE)

sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM") %>%  filter(strain != "LEW")
mean.all.other <- mean(sans_high_female$mean_co, na.rm=TRUE)

G.female.diff <- (mean.G.female / mean.all.other)
LEW.female.diff <- (mean.LEW.female / mean.all.other)
MSM.female.diff <- (mean.MSM.female / mean.all.other)
```

```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=FALSE}
Musc.female_mouse.av <-  female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc")

#variance plots
musc.female.VAR.plot <- ggplot(Musc.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
musc.female.VAR.plot

#Mol.female.VAR.plot <- ggplot(Mol.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
#Mol.female.VAR.plot

Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Dom feale ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
Dom.female.VAR.plot

```{r Female.specific.glm, echo=FALSE}

glm.female.strain.subsp <- glm(mean_co ~ subsp * strain, data= female_mouse.avs_for.analysis)

summry.female.strain.effects <- summary(glm.female.strain.subsp)
#round(summry.female.strain.effects$coefficients[,c(1,4)],5)
#write.table(summry.female.strain.effects$coefficients, "~./MLH1repo/summry.female.strain.effects.csv", sep=",",row.names = TRUE)

MLH1.female.spc.glm.M2 <- glm(mean_co ~ strain, data= female_mouse.avs_for.analysis)
sum_MLH1.female.spc.glm.M2 <- summary(MLH1.female.spc.glm.M2)

#round(lm.female.strain_sum$coefficients[,c(1,4)],5)

#write.table(lm.female.strain_sum$coefficients, "~./MLH1repo/lm.female.strain.csv", sep=",",row.names = TRUE)
```

```{r FEMALE.VAR.MODELS, echo=FALSE}
## within mouse variance models 
lmVAR.female.strain.subsp <- glm(var ~ subsp * strain, data= female_mouse.avs_for.analysis)
summry.female.strain.VAReffects <- summary(lmVAR.female.strain.subsp)

lmVAR.female.strain <- glm(var ~ strain, data= female_mouse.avs_for.analysis)
lmVAR.female.strain_sum <- summary(lmVAR.female.strain)

#round( #round isn't working for this sumary
#lmVAR.female.strain_sum$coefficients[,c(1,4)]
#Q12
Q12_female_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "female")

lm.m1.Q12_female <- glm(var ~ subsp * strain, data= Q12_female_table)
lm.m1.Q12_female_sum <- summary(lm.m1.Q12_female)

lm.m2.Q12_female <- glm(var ~ strain, data= Q12_female_table)
lm.m2.Q12VAR.female_sum <- summary(lm.m2.Q12_female)#LEW changes to PWD

lm.m2.Q12_female.cv <- glm(cV ~ strain, data= Q12_female_table)
lm.m2.Q12_female.cv_sum <- summary(lm.m2.Q12_female.cv)

#lm.m2.Q12VAR.female_sum$coefficients[,c(1,4)]
#LEW sig

```
<!-- male section -->

@@

```{r male.diff, echo=FALSE}
just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")

#how much greater are the High Rec gwRRs
just.male_mouse.av$Rec.group <- ifelse(grepl("PWD", just.male_mouse.av$strain), 1, 
                              ifelse(grepl("MSM", just.male_mouse.av$strain), 1,
                                ifelse(grepl("SKIVE", just.male_mouse.av$strain), 1, 0)))

#how much bigger are the high rec male means?
sans_LewRec_male <- just.male_mouse.av %>% filter(Rec.group == 0)
#mean(sans_LewRec_male$mean_co)

PWD_male <-  just.male_mouse.av %>% filter(strain == "PWD")
#mean(PWD_male$mean_co) / mean(sans_LewRec_male$mean_co)

MSM_male <- just.male_mouse.av %>% filter(strain == "MSM")
#mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co)

SKIVE_male <-   just.male_mouse.av %>% filter(strain == "SKIVE")
#mean(SKIVE_male$mean_co) /  
```

@@

```{r Male.specific.glm, echo=FALSE}
# 1 way anova, is at least 1 strain different from the rest?
# given a lm, how sig is the strain effect
glm.male.strain.subsp <- glm(mean_co ~ subsp * strain, data= just.male_mouse.av)

#summry.male.strain.effects <- summary(glm.male.strain.subsp)
#round(summry.male.strain.effects$coefficients[,c(1,4)], 5)

#write.table(summry.male.strain.effects$coefficients , "~./MLH1repo/summry.male.strain.effects.csv", sep=",",row.names = TRUE)

#1. determine which strain effects are sig

#M2
MLH1.male.specific.glm.M2 <- glm(mean_co ~ strain, data= just.male_mouse.av)
sum_MLH1.male.specific.glm.M2 <- summary(MLH1.male.specific.glm.M2)

#round(lm.male.strain_sum$coefficients[,c(1,4)], 5)

#write.table(lm.male.strain_sum$coefficients , "~./MLH1repo/lm.male.strain_sum.csv", sep=",",row.names = TRUE)

#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
```

```{r MALE.SPECIFIC.variance.model, echo=FALSE}
## within mouse variance models 
lmVAR.male.strain.subsp <- lm(var ~ subsp * strain, data= just.male_mouse.av)
summry.male.strain.VAReffects <- summary(lmVAR.male.strain.subsp)
#round(summry.male.strain.VAReffects$coefficients[,c(1,4)], 5)

#all p values kinda high, MSM and TOM are slightly sig
lmVAR.male.strain <- lm(var ~ strain, data= just.male_mouse.av)
lmVAR.male.strain_sum <- summary(lmVAR.male.strain)
#round(lmVAR.male.strain_sum$coefficients[,c(1,4)], 5)
#non sig

lmcv.male.strain <- lm(cV ~ strain, data= just.male_mouse.av)
lmcv.male.strain_sum <- summary(lmcv.male.strain)
#round(lmcv.male.strain_sum$coefficients[,c(1,4)], 5)
#HMI sig for cv model -- in the other model, CAST subsp and TOM strain are slightly significant

#Q12
Q12_male_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "male")

lm.m1.Q12_male <- lm(var ~ subsp * strain, data= Q12_male_table)
lm.m1.Q12_male_sum <- summary(lm.m1.Q12_male)
#round(lm.m1.Q12_male_sum$coefficients[,c(1,4)], 5)
#NS

lm.m2.Q12_male <- lm(var ~ strain, data= Q12_male_table)
lm.m2.Q12VAR.male_sum <- summary(lm.m2.Q12_male)
#round(lm.m2.Q12VAR.male_sum$coefficients[,c(1,4)], 5)
#NS

@@

```{r MLH1co.lmer.MM, echo=FALSE, waring=FALSE, message=FALSE}

#DF.HetC.MixedModel.HQ

MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
MLH1.MM.M1.reduce.sex <- lmer( mean_co ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1.MM.M1.reduce.subp <- lmer( mean_co ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )

#summary(MLH1.MM.M1)$coefficients

MLH1.MM_interaction <- drop1(MLH1.MM.M1, test="Chisq")$`Pr(Chi)`[2]
MLH1.MM_sex <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.sex)$`Pr(>Chisq)`[2]
MLH1.MM_subp <- anova(MLH1.MM.M1, MLH1.MM.M1.reduce.subp)$`Pr(>Chisq)`[2]
MLH1.MM_strain<- MM.rand.pval <- exactRLRT(MLH1.MM.M1)$p.value

MLH1.MMlmer_results <- data.frame(sub.sp = MLH1.MM_subp, 
sex.results =  MLH1.MM_sex, 
               interaction.pval = MLH1.MM_interaction,  rand.pval = 
MLH1.MM_strain)

```{r MLH1.glmM2, echo=FALSE}

remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
#kkll <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]

#strain = fixed
MLH1.M2.glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]   )

summ.MLH1.M2.glm.strain.fixed.full <- summary(MLH1.M2.glm.strain.fixed.full)

#table of just coeff anf pvalue
#round(summ$coefficients[,c(1,4)],5 )

#write.table(summ$coefficients , "~./MLH1repo/glm.strain.fixed.full.csv", sep=",",row.names = TRUE)
```

```{r MLH1.glm.M3, echo=FALSE}

remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")

MLH1.M3.strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ[  ! DF.HetC.MixedModel.HQ$strain %in% remove_strains,])

summ.MLH1.M3.strain.fixed <- summary(MLH1.M3.strain.fixed)

#round(summ.strain.fixed$coefficients[,c(1,4)],5)

#write.table(summ.strain.fixed$coefficients , "~./MLH1repo/summ.strain.fixed.csv", sep=",",row.names = TRUE)
```

```{r bret.comment.var, echo=FALSE}

@@

# Within mouse variance in CO count per cell

@@

```{r M1.MLH1.var_cv, echo=FALSE, warning=FALSE, message=FALSE}

MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )

@@

```{r M2.MLH1.var, echo=FALSE, include=FALSE}
#not including this model for now / brivity
#$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$

#why are these named M3? -- rename to M2

MLH1var_M2 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )

MLH1var_M2_sum <- summary(MLH1var_M2)

#summary(MLH1var_M2)
#anova(MLH1var_M2)  
#as.data.frame( round(summary(MLH1var_M2)$coefficients, 6)) #table of p values for coefficients.

MLH1cv_M2 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)

#summary(MLH1cv_M3)

#anova(MLH1cv_M2)
#as.data.frame( round(summary(MLH1cv_M2)$coefficients, 6))

#Q12 var
MLH1var_M2_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )

#summary(MLH1var_M3_Q12)
#anova(MLH1var_M2_Q12)  
#as.data.frame( round(summary(MLH1var_M2_Q12)$coefficients, 6))

@@

#Q12 cv
MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
#round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
```

@@

# Q1 Sex Differences in the Length of the Synaptonemal Complex

@@

#TOTAL SC AREA!!
lmer.Q1_mus.av_totSC <- lmer(mean.skel ~ subsp * sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)

@@

#Q1.lmer.totSC_results

#G
```

@@

```{r Q2.SC.length.bret.comments, echo=FALSE}
#correlation, total sc and Mean MLH1

@@

Q2.glm_mus.av_totSC_M1 <- glm(mean.skel ~ subsp * strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
sum.Q2.glm_mus.av_totSC_M1  <- summary(Q2.glm_mus.av_totSC_M1)$coefficients

#sum.Q2.glm_mus.av_totSC_M1$coefficients[,4]   pvalues of all effects
#sum.Q2.glm_mus.av_totSC_M1$coefficients[,1]  coefficients of all effects
#round(sum.Q2.glm_mus.av_totSC_M1$coefficients[,c(1,4)], 6)

Q2.glm_mus.av_totSC_M2 <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
sum.Q2.glm_mus.av_totSC_M2  <- summary(Q2.glm_mus.av_totSC_M2)
#round(summary(sum.Q2.glm_mus.av_totSC_M2)$coefficients[,c(1,4)], 6)

@@

#Dom
Q2.glm_mus.av_totSC_M2_DOM <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom"))
sum.Q2.glm_mus.av_totSC_M2_DOM  <- summary(Q2.glm_mus.av_totSC_M2_DOM)
#there are significant strain effects in DOM,

#Musc
Q2.glm_mus.av_totSC_M2_MUSC <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )
sum.Q2.glm_mus.av_totSC_M2_MUSC  <- summary(Q2.glm_mus.av_totSC_M2_MUSC)
#KAZ -- is the only significant effect

#Mol
Q2.glm_mus.av_totSC_M2_MOL <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Mol") )
sum.Q2.glm_mus.av_totSC_M2_MOL  <- summary(Q2.glm_mus.av_totSC_M2_MOL)

#these pair-wise t.test aren't working
#dom
#pairwise.t.test( ( totSC.mouse.av.table %>% filter(subsp == "Dom") %>% filter(sex == "male") )$mean.skel, 
#                 ( totSC.mouse.av.table %>% filter(subsp == "Dom") %>% filter(sex == "male") )$strain )

#musc
#pairwise.t.test( ( totSC.mouse.av.table %>% filter(subsp == "Musc") %>% filter(sex == "male") )$mean.skel, 
#                 ( totSC.mouse.av.table %>% filter(subsp == "Musc") %>% filter(sex == "male") )$strain )

#mol
#pairwise.t.test( ( totSC.mouse.av.table %>% filter(subsp == "Mol") %>% filter(sex == "male") )$mean.skel, 
#                 ( totSC.mouse.av.table %>% filter(subsp == "Mol") %>% filter(sex == "male") )$strain )

@@

#Q2.glm.M2.SC.cell.av_long.biv <- glm(mean.SC ~ strain, data= cell.av.long.biv.table)
#round(summary(glm.Q2.M2.SC.cell.av_long.biv)$coefficients[,c(1,4)], 6)
#table(mouse.av.long.biv.table$mouse)
#Q2mouse.av.long.biv.table

#this dataset is all males
glmQ2_long.biv_M1 <- glm(mean.SC ~ subsp * strain, data= Q2mouse.av.long.biv.table)
sum.glmQ2_long.biv_M1 <- summary(glmQ2_long.biv_M1)
#

glmQ2_long.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.long.biv.table)
sum.glmQ2_long.biv_M2 <- summary(glmQ2_long.biv_M2)
#round(sum.glmQ2_long.biv_M2$coefficients[,c(1,4)], 6)

glmQ2_long.biv_M2_DOM <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )  )
sum.glmQ2_long.biv_M2_DOM <- summary(glmQ2_long.biv_M2_DOM)

#MUSC
glmQ2_long.biv_M2_MUSC <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )  )
sum.glmQ2_long.biv_M2_MUSC <- summary(glmQ2_long.biv_M2_MUSC)

#MOL can't be done -- only 1 mouse for each MSM and MOLF
glmQ2_long.biv_M2_MOL <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Mol") )  )
sum.glmQ2_long.biv_M2_MOL <- summary(glmQ2_long.biv_M2_MOL)

@@

```{r SC.length.male_tests, echo=FALSE, include=FALSE, eval=FALSE}
## run the tests to check following expectations

#1.Dom strains not different from ech other
#2. PWD and SKIVE are different from the other strains
#3.MSM and MOLF are different from each other

#male.Musc.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"), ]

male.Dom.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom"), ]
#the subsp labels get messed up, unless they are designated chr first
Dom.male_SC.length_lm <- lm(mean_SC~ strain,  data =male.Dom.mouse.av  )

Dom.male_SC.length_lm.1 <- lm(mean.SC_1CO~ strain/mouse,  data =male.Dom.mouse.av  ) #wierd sytax 
Dom.male_SC.length_lm.2 <- lm(mean.SC_2CO~ strain,  data =male.Dom.mouse.av  )
#Dom.male_SC.length_lm.3 <- lm(mean.SC_3CO~ strain,  data =male.Dom.mouse.av  ) #only WSB and LEW?
#total sc sig, 1CO sig, 2CO sig

Dom.male_SC.aov <- aov(mean_SC ~ strain, data = male.Dom.mouse.av)
#0.000369 ***
Dom.male_1CO_SC.aov <- aov(mean.SC_1CO ~ strain, data = male.Dom.mouse.av)
#0.000816 ***
Dom.male_2CO_SC.aov <- aov(mean.SC_2CO ~ strain, data = male.Dom.mouse.av)
#0.0148 *

show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()

show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean.SC_2CO, x= strain, color=strain))+ geom_jitter()
#making plots above show that mouse average SC lengths are lower could there be microscope issue?

#make sure to remove some of the juvi mice!! from the AP_mouse table
show.Dom.male.CO <- ggplot(
male_mouse.av[(male_mouse.av$subsp.x == "Dom"),], aes(y= mean_co, x= strain.x, color=strain.x))+ geom_jitter()

Dom.male_CO.aov <- aov(mean_co ~ strain, data = male_mouse.av[ (male_mouse.av$subsp == "Dom"),])
#P not sig for CO mean

#seems like there are more mice in the MLH1 data set than bivData
#

#WSB all dissected and imaged in '15
#G imaged 15 - 16
#LEW imaged 15 to 17

ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          ylab = "", xlab = "")

show.Dom.male.SC <- ggplot(male.Dom.mouse.av, aes(y= mean_SC, x= strain, color=strain))+ geom_jitter()

ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean_SC", 
          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          ylab = "", xlab = "")

ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_1CO", 
          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          ylab = "1CO SC Length", xlab = "")

ggboxplot(data = male.Dom.mouse.av, x = "strain", y = "mean.SC_2CO", 
          color = "strain", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          ylab = "2CO SC Length", xlab = "")
```

```{r SC.length.pairwise.T, echo=FALSE, include=FALSE, eval=FALSE}

#REMOVE RUNS COMPARISONS ACROSS FULL DATASET
#male.Dom.mouse.av, male.Mol.mouse.av, male.Musc.mouse.av

male.Musc.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc"),]
male.Dom.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom"),]
male.Mol.mouse.av <- Male.poly.Mouse.Table_BivData_4MM[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol"),]

Pwise.ttest_Dom.male <- pairwise.t.test(male.Dom.mouse.av$mean_SC, male.Dom.mouse.av$strain)
#G-WSB is sig, WSB-LEW are slightly sig

Pwise.ttest_Musc.male <- pairwise.t.test(male.Musc.mouse.av$mean_SC, male.Musc.mouse.av$strain)
#PWD and MSM are most diff from SKIVE?

Pwise.ttest_Mol.male <- pairwise.t.test(male.Mol.mouse.av$mean_SC, male.Mol.mouse.av$strain)
#the pooled SC value tests are sig

#the mol comparisons are different due to adj p values

Pwise.ttest_Dom.male.1CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_1CO, male.Dom.mouse.av$strain)
#G-WSB is sig, WSB-LEW are slightly sig

Pwise.ttest_Musc.male.1CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_1CO, male.Musc.mouse.av$strain)
#no sig (only 3) -- does include CZECCH

#non-sig, with mol pooled with musc

Pwise.ttest_Mol.male.1CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_1CO, male.Mol.mouse.av$strain)

Pwise.ttest_Dom.male.2CO <- pairwise.t.test(male.Dom.mouse.av$mean.SC_2CO, male.Dom.mouse.av$strain)

Pwise.ttest_Musc.male.2CO <- pairwise.t.test(male.Musc.mouse.av$mean.SC_2CO, male.Musc.mouse.av$strain)
#non of these sig with the mulitple adjust

Pwise.ttest_Mol.male.2CO <- pairwise.t.test(male.Mol.mouse.av$mean.SC_2CO, male.Mol.mouse.av$strain)

#pairwise.t.test(Ozone, Month, p.adj = "bonf")
```

@@

<!-- t tests mouse.avs_4MM, REAL.short_mouse_table-->

```{r Q2.1COpos_mouse.av.t.test_all, echo=FALSE}
#mouse.avs_4MM, #31may19_SKIVE_m2  has NA for norm.first.foci.pos

#High = MSM, PWD, SKIVE
Q2.1COpos_Ttest <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "0"))$norm.first.foci.pos, 
            (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group == "1"))$norm.first.foci.pos )
#p-value = 1

#high = PWD and MSM
Q2.1COpos_Ttest.HL.ex <- t.test( (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group.exclu == "0"))$norm.first.foci.pos, 
            (mouse.avs_4MM %>% filter(sex == "male") %>% filter(Rec.group.exclu == "1"))$norm.first.foci.pos )
#p-value = 1

```{r Q2.1COpos_within.subsp_ttest, echo=FALSE}

#Dom (predicted no difference) -- can't DO a t.test for DOM across groups..

#Musc
Q2.1COpos_mus.av_Musc_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Musc") %>% filter(Rec.group == "0"))$norm.first.foci.pos,

            (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Musc") %>% filter(Rec.group == "1"))$norm.first.foci.pos )
# p-value = 0.4

#Mol
Q2.1COpos_mus.av_Mol_ttest  <-  t.test( (mouse.avs_4MM %>%  filter(sex == "male") %>%  filter(subsp == "Mol") %>% filter(Rec.group == "0"))$norm.first.foci.pos,

              (mouse.avs_4MM  %>%  filter(sex == "male") %>% filter(subsp == "Mol") %>% filter(Rec.group == "1"))$norm.first.foci.pos   )
# p-value = 0.07
# exclusive rec group not sig
```

```{r Q2.1CO_mouse.av.short.long_t.test, echo=FALSE}
#these don't have mean.first.foci

#short -- real.long.bivData -- REAL.short_mouse_table
Q2.1COposABS_short_ttest_HL <- t.test( (REAL.short_mouse_table %>% filter(sex == "male") %>%  filter(Rec.group == "0"))$mean_F1pos_ABS,

      (REAL.short_mouse_table %>% filter(sex == "male")  %>% filter(Rec.group == "1"))$mean_F1pos_ABS   )
#p-value = 0.3

Q2.1COpos.PER_short_ttest_HL <- t.test( (REAL.short_mouse_table %>% filter(sex == "male") %>% filter(Rec.group == "0"))$mean_F1pos_PER, 
      (REAL.short_mouse_table %>% filter(sex == "male")  %>% filter(Rec.group == "1"))$mean_F1pos_PER   )
# p-value = 0.1

#remove SKIVE??
```

```{r Q2.1CO.pos_short.long.ttest, echo=FALSE}

#real.long.bivData
#single biv level
Rec.group_LONG.BivData_F1post_Ttest <- t.test(
  (real.long.bivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "0") )$PER_Foci_1, 
(real.long.bivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "1"))$PER_Foci_1  )
#p-value = 0.5

Rec.group_BivData_F1post_Ttest <- t.test(
  (Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "0") )$PER_Foci_1, 
(Curated_BivData %>% filter(sex == "male") %>% filter(hand.foci.count == 1) %>% filter(Rec.group == "1"))$PER_Foci_1  )
#p-value = 0.2

>>>>

#write the table out

@@

```{r mouse.var.plots.female, echo=FALSE, fig.width = 4, fig.height=2.5, include=FALSE, eval=FALSE}

#variance plots
musc.female.VAR.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(sex== "female") %>% filter(subsp == "Musc") ) , aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y =var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Musc female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
musc.female.VAR.plot

Mol.female.VAR.plot <- ggplot( (female_mouse.avs_for.analysis %>% filter(subsp == "Mol") ), aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("Mol female of within mouse MLH1 distributions")+scale_color_manual(values=colors_of_strains)
Mol.female.VAR.plot

Dom.female.VAR.plot <- ggplot(Dom.female_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain), width = .2 )+ ggtitle("Dom female\nwithin mouse  MLH1 distributions") + scale_color_manual(values=colors_of_strains)
Dom.female.VAR.plot

<!-- male section -->

@@

@@

@@

@@

# Within mouse **variance** in CO count per cell

@@

```{r M1.MLH1.var_cv, echo=FALSE, warning=FALSE, message=FALSE}
MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1var_M1.reduce.sex <- lmer( var ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1var_M1.reduce.subp <- lmer( var ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )

@@

```{r M2.MLH1.var, echo=FALSE, include=FALSE}

MLH1var_M2 <- lm( var ~ subsp * sex * strain,  data=DF.HetC.MixedModel.HQ )
MLH1var_M2_sum <- summary(MLH1var_M2)

#anova(MLH1var_M2)  
#as.data.frame( round(summary(MLH1var_M2)$coefficients, 6)) #table of p values for coefficients.

MLH1cv_M2 <- lm(cV ~ subsp*sex*strain, data=DF.HetC.MixedModel.HQ)
#summary(MLH1cv_M2)
#anova(MLH1cv_M2)
#as.data.frame( round(summary(MLH1cv_M2)$coefficients, 6))

#Q12 var
MLH1var_M2_Q12 <- lm( var ~ subsp*sex*strain,  data=Q12_mouse_table_w.Ages )
#summary(MLH1var_M3_Q12)
#anova(MLH1var_M2_Q12)  
#as.data.frame( round(summary(MLH1var_M2_Q12)$coefficients, 6))

@@

#Q12 cv
#MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
#round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
```

@@

# Q1. Sex Differences in the Length of the Synaptonemal Complex

@@

#TOTAL SC AREA
lmer.Q1_mus.av_totSC <- lmer(mean.skel ~ subsp * sex + (1|strain), data=Q1.lmer_totSC.mouse.av.table, REML = TRUE)

@@

@@

@@

Q2.glm_mus.av_totSC_M4 <- glm(mean.skel ~ subsp * strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
sum.Q2.glm_mus.av_totSC_M4  <- summary(Q2.glm_mus.av_totSC_M1)$coefficients

#sum.Q2.glm_mus.av_totSC_M4$coefficients[,4]   pvalues of all effects
#sum.Q2.glm_mus.av_totSC_M4$coefficients[,1]  coefficients of all effects
#round(sum.Q2.glm_mus.av_totSC_M4$coefficients[,c(1,4)], 6)

#M4.2? (removed subspecies?)

Q2.glm_mus.av_totSC_M4.2 <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male"))
sum.Q2.glm_mus.av_totSC_M4.2  <- summary(Q2.glm_mus.av_totSC_M4.2)
#round(summary(sum.Q2.glm_mus.av_totSC_M2)$coefficients[,c(1,4)], 6)

@@

#Dom
Q2.glm_mus.av_totSC_M4.2_DOM <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Dom"))
sum.Q2.glm_mus.av_totSC_M4.2_DOM  <- summary(Q2.glm_mus.av_totSC_M4.2_DOM)
#there are significant strain effects in DOM,

#Musc
Q2.glm_mus.av_totSC_M4.2_MUSC <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Musc") )
sum.Q2.glm_mus.av_totSC_M4.2_MUSC  <- summary(Q2.glm_mus.av_totSC_M4.2_MUSC)
#KAZ -- is the only significant effect

#Mol
Q2.glm_mus.av_totSC_M4.2_MOL <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male") %>% filter(subsp == "Mol") )
sum.Q2.glm_mus.av_totSC_M4.2_MOL  <- summary(Q2.glm_mus.av_totSC_M4.2_MOL)

@@

#this dataset is all males
glmQ2_long.biv_M4 <- glm(mean.SC ~ subsp * strain, data= Q2mouse.av.long.biv.table)
sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)

glmQ2_long.biv_M4 <- glm(mean.SC ~ strain, data= Q2mouse.av.long.biv.table)
sum.glmQ2_long.biv_M4 <- summary(glmQ2_long.biv_M4)
#round(sum.glmQ2_long.biv_M2$coefficients[,c(1,4)], 6)

glmQ2_long.biv_M4.2_DOM <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Dom") )  )
sum.glmQ2_long.biv_M4.2_DOM <- summary(glmQ2_long.biv_M4.2_DOM)

#MUSC
glmQ2_long.biv_M4.2_MUSC <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Musc") )  )
sum.glmQ2_long.biv_M4.2_MUSC <- summary(glmQ2_long.biv_M4.2_MUSC)

#MOL can't be done -- only 1 mouse for each MSM and MOLF
glmQ2_long.biv_M4.2_MOL <- glm(mean.SC ~ strain, data= ( Q2mouse.av.long.biv.table %>% filter(subsp == "Mol") )  )
sum.glmQ2_long.biv_M4.2_MOL <- summary(glmQ2_long.biv_M4.2_MOL)

@@

@@
